COCT Urban Mobility

### 2023-08-13 draft_v0_1.Rmd
Dashboard mock-up added. From Windows.

@@

## 4. Identify value-adding insights for the management of Urban Mobility, from the dataset provided, in regard to commuter transport within the City.

@@

@@

 unique(dta_dir$code)

 code_txt <- "Conjested Traffic at Intersection"

@@

x <- sort(tmp$n, decreasing = TRUE)
x <- x[6]

tmp %>% 
mutate(day_of_week = wday(creation_timestamp, label=TRUE),
       label = case_when(n > x ~ 
                           paste(day_of_week, creation_timestamp),
                         TRUE ~ "")) %>%

 ggplot(aes(creation_timestamp, n, group = code,
            label = label))+
 geom_line(colour = "grey70")  +
 geom_point()+
   geom_text()

dta_dir %>% 
 mutate(create_date = as_date(creation_timestamp)) %>% 
 filter(code == code_txt) %>% 
 group_by(code,create_date) %>% 
 count() %>% 
 ungroup() %>% 
  filter(n > x) %>% 
  select(-n) %>% 
  inner_join(dta_dir) %>% 
  group_by(create_date,official_suburb) %>% 
  count() %>% 
  ungroup() %>% 
  ggplot(aes(n, reorder(official_suburb,n), fill = as.factor(create_date)))+
  geom_col()

@@

 dta_dir %>% 
   filter(code %in% x) %>% 
 mutate(day_of_week = wday(creation_timestamp, label=TRUE)) %>% 
 group_by(code, day_of_week) %>% 
 count() %>%
   ggplot(aes(day_of_week, n, group = code, colour = code))+
 geom_point()+
 geom_line()

 dta_dir %>% 
 mutate(day_of_week = wday(creation_timestamp, label=TRUE)) %>% 
 filter(code == code_txt) %>% 
 group_by(day_of_week) %>% 
 count() %>% 
 ggplot(aes(day_of_week, n))+
 geom_col()

 # ?stats
 # odds of having congestion on a wednesday vs other weekdays

 # suburbs and days of week

@@

sz <- 3
 dta_dir %>% 
 mutate(day_of_week = wday(creation_timestamp, label=TRUE)) %>% 
 group_by(day_of_week, official_suburb) %>% 
 count() %>% 
   mutate(label = case_when(n >= x ~ official_suburb,
                          TRUE ~ "")) %>% 
   ggplot(aes(day_of_week, n, label = label))+
   geom_point(alpha = 0.6)+
   geom_text(alpha = 0.5, size = sz, check_overlap = TRUE)

@@

>>>>

# fix typo in code

sort(unique(dta$code[dta$directorate == "URBAN MOBILITY"]))
dta$code <- gsub("Conjested", "Congested", dta$code) 
```

@@

theme_tmp <- theme_doc+
  theme(text = element_text(size = 9))
theme_set(theme_tmp)
# top suburbs

tmp <- dta_dir %>% 
  group_by(official_suburb) %>% 
  count() %>% 
  ungroup() %>% 
  mutate(percentage = (n/x*100)) %>% 
  arrange(desc(n)) %>% 
  slice(1:5)

p1 <- tmp %>% 
  ggplot(aes(percentage, reorder(official_suburb, percentage),
             fill = official_suburb))+
  geom_col()+
  scale_fill_manual(values = coct_cols)+
  labs(title = "Top 5 Suburbs Contributing Requests")+
  theme(legend.title = element_blank(),
        axis.title.y = element_blank())

top5 <- tmp$official_suburb

p2 <- dta_dir %>% 
  group_by(month_creation) %>% 
  count(name = "total_month") %>% 
  inner_join(dta_dir %>% 
  group_by(month_creation, official_suburb) %>% 
  count()) %>% 
  ungroup() %>% 
  filter(official_suburb %in% top5) %>% 
  mutate(percentage = (n/total_month*100)) %>% 
  arrange(desc(percentage)) %>% 
  ggplot(aes(month_creation, percentage, colour = official_suburb,
             group = official_suburb))+
  geom_point(size = 3)+
  geom_line(linewidth = 2)+
  theme(legend.title = element_blank(),
        axis.title.y = element_blank())+
  scale_colour_manual(values = coct_cols)+
  labs(title = "Top 5 Suburbs Contributing Requests, Per Month")

# top codes for top 5 suburbs

tmp <- dta_dir %>% 
  filter(official_suburb %in% top5) %>% 
  group_by(month_creation, official_suburb) %>% 
  count(name = "total_month") %>% 
  ungroup() %>% 
  inner_join(dta_dir %>% 
  filter(official_suburb %in% top5) %>% 
  group_by(month_creation, official_suburb, code) %>% 
  count() %>% 
  ungroup() %>% 
  filter(n >= n_cutoff)) %>% 
  mutate(percentage = n/total_month*100)

p3 <- tmp %>% 
  ggplot(aes(percentage, reorder(code, percentage),
             fill = month_creation))+
  geom_col(position = "dodge")+
  facet_wrap(~official_suburb, scales = "free_y",
             ncol = 1)+
  scale_fill_manual(values = coct_cols)+
  theme(legend.title = element_blank(),
        axis.title.y = element_blank())

A <- ggpubr::ggarrange(p1, p2, common.legend = TRUE, ncol =1)
A <- ggpubr::ggarrange(A, p3, nrow = 1)

# completion times

  # flag things from past three months with missing completion times
  # create dataset as if it is 1 April 2020

dta_apr <- dta_dir %>% 
  mutate(date_complete = as_date(completion_timestamp),
         date_create = as_date(creation_timestamp),
        across(c(completion_timestamp, date_complete, month_completion),~ifelse(month_completion %in% c("jan", "feb", "mar"),
                                .x, NA)),
        date_today = as_date("2020-04-01"))

tmp <- dta_apr  %>% 
  filter(is.na(completion_timestamp)) %>% 
  mutate(interv = date_create %--% date_today,
          time_elapsed_weeks = as.duration(interv) / dweeks(1)) %>% 
  select(-interv) %>% 
  filter(time_elapsed_weeks > 4)

p1 <- tmp %>% 
  group_by(code) %>% 
  count() %>% 
  ungroup() %>% 
  arrange(desc(n)) %>% 
  slice(1:5) %>% 
  ggplot(aes(n, reorder(code, n)))+
  geom_col(fill = coct_cols[2])+
  theme(axis.title.y  = element_blank())+
  labs(title = "Top 5 Codes Pending Requests for 4 Weeks or More")

p2 <- tmp %>% 
  group_by(official_suburb) %>% 
  count() %>% 
  ungroup() %>% 
  arrange(desc(n)) %>% 
  slice(1:5) %>% 
  ggplot(aes(n, reorder(official_suburb, n)))+
  geom_col(fill = coct_cols[1])+
  theme(axis.title.y  = element_blank())+
  labs(title = "Top 5 Suburbs Pending Requests for 4 Weeks or More")

B <- ggpubr::ggarrange(p1, p2, ncol = 1)

# city-wide completion time med, q80

p1 <- time_dta_list$all_times %>% 
  filter(n >= n_cutoff & official_suburb == sub_txt &
           code != code_txt) %>% 
  pivot_longer(c(med_time_hours, q80_time_hours),
               names_to = "quantile", 
               values_to = "hours") %>% 
  ggplot(aes(log10(hours), reorder(code,hours), group = code, colour = quantile))+
   geom_line(colour = "grey70")+
  geom_point(size = 3)+
   scale_colour_manual(values = coct_cols)+
  theme(axis.title.y = element_blank())

p2 <- time_dta_list$all_times %>% 
  filter(n >= n_cutoff & official_suburb != sub_txt &
           code != code_txt) %>% 
  arrange(desc(med_time_hours)) %>% 
  slice(1:5) %>% 
  pivot_longer(c(med_time_hours, q80_time_hours),
               names_to = "quantile", 
               values_to = "hours") %>% 
  ggplot(aes(log10(hours), reorder(code,hours), group = code, colour = quantile))+
   geom_line(colour = "grey70")+
  geom_point(size = 3)+
   scale_colour_manual(values = coct_cols)+
  theme(axis.title.y = element_blank())+
  facet_wrap(~reorder(official_suburb,hours), scale = "free_y", ncol = 1)

C <- ggpubr::ggarrange(p1, p2, nrow = 1)

D <- ggpubr::ggarrange(A,  B, labels = c("A", "B"))
E <- ggpubr::ggarrange(C, labels = "C")

ggpubr::ggarrange(D, E, ncol = 1)

theme_set(theme_doc)

knit_exit()

@@

Traffic congestion can be explored by plotting request numbers associated with this code, over time, and looking for patterns in day of week, or highlighting particularly bad days (Fig. x). Several days in February 2020 stand out.

@@

 code_txt <- "Congested Traffic at Intersection"

@@

x <- sort(tmp$n, decreasing = TRUE)
x <- x[6] # for top 5

tmp %>% 
mutate(day_of_week = wday(creation_timestamp, label=TRUE),
       label = case_when(n > x ~ 
                           paste(day_of_week, creation_timestamp),
                         TRUE ~ ""),
       weekend = case_when(day_of_week %in% c("Sat", "Sun") ~ "weekend",
                           TRUE ~ "week")) %>%

 ggplot(aes(creation_timestamp, n, group = code,
            label = label, colour = weekend))+
 geom_line(colour = "grey70")  +
 geom_point(size = 3)+
   geom_text()+
  scale_color_manual(values = coct_cols)+
  theme(legend.title = element_blank())

The suburbs contributing to the top 5 days with highest number of traffic congestion requests, show that a small number of suburbs make the largest contribution.

```{r fig.height = 16}
dta_dir %>% 
 mutate(create_date = as_date(creation_timestamp)) %>% 
 filter(code == code_txt) %>% 
 group_by(code,create_date) %>% 
 count() %>% 
 ungroup() %>% 
  filter(n > x) %>% 
  select(-n) %>% 
  inner_join(dta_dir) %>% 
  group_by(create_date,official_suburb) %>% 
  count() %>% 
  ungroup() %>% 
  ggplot(aes(n, reorder(official_suburb,n), fill = as.factor(create_date)))+
  geom_col()+
  scale_fill_manual(values = coct_cols)+
  labs(fill = "Top 5 Dates")+
  theme(axis.title.y = element_blank())

# get outlier date
tmp <- dta_dir %>% 
  mutate(date = as_date(creation_timestamp)) %>% 
  filter(code == code_txt & official_suburb == "PARKLANDS") %>% 
  group_by(date) %>% 
  count() %>% 
  ungroup() %>% 
  arrange(desc(n)) %>% 
  slice(1) %>% 
  pull(date)

A clear pattern for number of requests and day of the week, is seen for congested traffic requests, and to some degree for Traffic lights out. Wednesdays contributed the most traffic congestion requests, but when removing the outlier for Parklands on `r tmp`,there seems to be a mid-week plateau, tapering down to Sunday and Saturday on either side, as expected.

@@

p1 <- dta_dir %>% 
  mutate(date = as_date(creation_timestamp)) %>% 
   filter(code %in% x ) %>% 
 mutate(day_of_week = wday(creation_timestamp, label=TRUE)) %>% 
 group_by(code, day_of_week) %>% 
 count() %>%
   ggplot(aes(day_of_week, n, group = code, colour = code))+
 geom_point(size = 3)+
 geom_line()+
   scale_colour_manual(values = coct_cols)+
  labs(title = "All values")

p2 <- dta_dir %>% 
  mutate(date = as_date(creation_timestamp)) %>% 
  filter(!(date == tmp & code == code_txt)) %>% 
   filter(code %in% x ) %>% 
 mutate(day_of_week = wday(creation_timestamp, label=TRUE)) %>% 
 group_by(code, day_of_week) %>% 
 count() %>%
   ggplot(aes(day_of_week, n, group = code, colour = code))+
 geom_point(size = 3)+
 geom_line()+
   scale_colour_manual(values = coct_cols)+
  labs(title = "Parklands outlier removed")

ggpubr::ggarrange(p1, p2, common.legend = TRUE)

 # 
 # dta_dir %>% 
 # mutate(day_of_week = wday(creation_timestamp, label=TRUE)) %>% 
 # filter(code == code_txt) %>% 
 # group_by(day_of_week) %>% 
 # count() %>% 
 # ggplot(aes(day_of_week, n))+
 # geom_col()

The pattern is less pronounced when looking at individual suburbs, but the three suburbs identified previously for intervention, feature prominently in the highest numbers of requests(Fig x).

```{r}
 # suburbs and days of week

@@

sz <- 3
 dta_dir %>% 
 mutate(day_of_week = wday(creation_timestamp, label=TRUE)) %>% 
 group_by(day_of_week, official_suburb) %>% 
 count() %>% 
   mutate(label = case_when(n >= x ~ official_suburb,
                          TRUE ~ "")) %>%
   ggplot(aes(day_of_week, n, label = label))+
   geom_jitter(alpha = 0.6, width = 0.1)+
   ggrepel::geom_text_repel(alpha = 0.5, size = sz,
                            max.overlaps = 50,
                            force = 0.1)

@@

## Conclusion
